---
title: "Ovarian cancer, Cisplatin"
author: "H Qin"
date: "January 11, 2016"
output: html_document
---

Analysis of ovarian cancer under treatment of cisplatin
GEO Data: http://www.ncbi.nlm.nih.gov/geo/geo2r/?acc=GSE47856 
Use limma package for differnetial expression analysis. 
```{r message=FALSE}
#   Differential expression analysis with limma
rm(list=ls())
setwd("~/github/MJ_RISE2016/project2.GSE47856")
library(Biobase)
library(GEOquery)
library(limma)
```

Load series and platform data from GEO. After the first run, save the image for working offline. 
```{r}
# gset <- getGEO("GSE47856", GSEMatrix =TRUE)
# save.image("GSE47856.RData")
# rm(list=ls())
```

Load the image. 
```{r, comment=NA}
load("GSE47856.RData")
if (length(gset) > 1) idx <- grep("GPL6244", attr(gset, "names")) else idx <- 1
gset <- gset[[idx]]
# make proper column names to match toptable 
fvarLabels(gset) <- make.names(fvarLabels(gset))
```

Study the structure of gset
```{r, comment=NA}
# study the meta data
experimental_design = gset@phenoData@data
# gset@phenoData@varMetadata
# experimental_design[, "source_name_ch1"][1:10]
experimental_design[1:10,  c("characteristics_ch1", "characteristics_ch1.2")]
unique( experimental_design$characteristics_ch1 )
```

Pick one cell line for test. 
```{r}
sel = grep("HeyC2", experimental_design$characteristics_ch1)
gset = gset[ ,sel]
```

Do a histogram of the expression data
```{r}
ex <- exprs(gset)
hist(ex[,1])
hist(log2(ex[,1]))
boxplot(ex)
```


Log2 transform is unnecessary?  Not sure about LogC
```{r}
qx <- as.numeric(quantile(ex, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))
LogC <- (qx[5] > 100) ||
  (qx[6]-qx[1] > 50 && qx[2] > 0) ||
  (qx[2] > 0 && qx[2] < 1 && qx[4] > 1 && qx[4] < 2)
# if (LogC) { ex[which(ex <= 0)] <- NaN
# exprs(gset) <- log2(ex) }
# ex = log2(ex)
```

Are the microarray data normalized? The sum of each array are similar. 
```{r, comment=NA}
head(ex)
sums = apply(ex,2, sum)
sums / max(sums)
medians = apply(ex, 2, median)
medians
means = apply(ex, 2, mean)
means

```


Set up the data and proceed with analysis. Modified by GEO sample script. 
```{r, comment=NA}
gset@phenoData@data[,  c("characteristics_ch1", "characteristics_ch1.2")]
gset@phenoData@data$characteristics_ch1.2
treatments <- make.names(gset@phenoData@data$characteristics_ch1.2)
fl.names = unique(treatments)
fl = as.factor( LETTERS[1:length(fl.names)]) 
names(fl) = fl.names
gset$description <- fl
design <- model.matrix(~ description + 0, gset)
colnames(design) <- levels(fl)
fit <- lmFit(gset, design)
cont.matrix <- makeContrasts( A-B, levels=fl)
fit2 <- contrasts.fit(fit, cont.matrix)
fit2 <- eBayes(fit2, 0.01)
tT <- topTable(fit2, adjust="fdr", sort.by="B", number=50)
tT <- topTable(fit2, adjust="fdr", sort.by="B", number=length(fit2))
```


```{r}
str(fit)
str(fit2)
str(tT)
```

