---
title: "Maya Jones RISE Research Spring 2016, Paclitaxel-resistence in ovarian cancer cell lines"
output: html_document
---
Data are from http://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE2627

Paclitazel:  https://en.wikipedia.org/wiki/Paclitaxel

The reference paper is at http://www.ncbi.nlm.nih.gov/pubmed/18068629


Load libraries
```{r, comment=NA, message=FALSE}
library(Biobase)
library(GEOquery)
library(limma)
```

Retrive gene expression data from NCBI GEO database. 

```{r, comment=NA}
gset <- getGEO("GSE2627", GSEMatrix =TRUE)
if (length(gset) > 1) idx <- grep("GPL2022", attr(gset, "names")) else idx <- 1
gset <- gset[[idx]]

# make proper column names to match toptable 
fvarLabels(gset) <- make.names(fvarLabels(gset))

```

Study the structure of gset and its meta-data
```{r, comment=NA}
# gset@phenoData@varMetadata  #this is empty
experimental_design = gset@phenoData@data
experimental_design
experimental_design[, "source_name_ch1"]
# experimental_design[1:10,  c("characteristics_ch1", "characteristics_ch1.2")]
# unique( experimental_design$characteristics_ch1 )
```

Do a histogram of the expression data
```{r}
ex <- exprs(gset)
mat = matrix( seq(1,length(ex[1,])), nrow=length(ex[1,])/2, ncol= 2 ); 
layout(mat);
for(i in 1:length(ex[1,]) ) { 
  hist(ex[,i], main=experimental_design$source_name_ch1[i]) 
}
layout(1)
boxplot(ex)
```
All the expression data are centered around zero? No need for log2 transformation. 

According to GSE2627 summary. These data are log2 transformed expression ratios and are normalized by scaled loess normalization. 
```{r}
apply(ex, 2, mean)
apply(ex, 2, median)
```


