---
title: "Maya Jones RISE Research Spring 2016, Paclitaxel-resistence in ovarian cancer cell lines"
output: html_document
---
Data are from http://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE2627. 
"Data was transformed into log base 2 expression ratios and normalized using scaled loess normalization.""

Paclitazel:  https://en.wikipedia.org/wiki/Paclitaxel

The reference paper is at http://www.ncbi.nlm.nih.gov/pubmed/18068629

Cy3 and Cy5
https://en.wikipedia.org/wiki/Cyanine
Cy3 fluoresces greenish yellow (~550 nm excitation, ~570 nm emission), while Cy5 is fluorescent in the red region (~650 excitation, 670 nm emission)

Load libraries
```{r, comment=NA, message=FALSE}
library(Biobase)
library(GEOquery)
library(limma)
```

Retrive gene expression data from NCBI GEO database. 

```{r, comment=NA}
#gset <- getGEO("GSE2627", GSEMatrix =TRUE)
#save.image("GSE2627.RData")
load("GSE2627.RData")
if (length(gset) > 1) idx <- grep("GPL2022", attr(gset, "names")) else idx <- 1
gset <- gset[[idx]]
str(gset)
# make proper column names to match toptable 
fvarLabels(gset) <- make.names(fvarLabels(gset))
```

Study the structure of gset and its meta-data
```{r, comment=NA}
# gset@phenoData@varMetadata  #this is empty
experimental_design = gset@phenoData@data
experimental_design
experimental_design[, "source_name_ch1"]
# experimental_design[1:10,  c("characteristics_ch1", "characteristics_ch1.2")]
# unique( experimental_design$characteristics_ch1 )
```

Do a histogram of the expression data
```{r}
ex <- exprs(gset)
# mat = matrix( seq(1,length(ex[1,])), nrow=length(ex[1,])/2, ncol= 2 ); 
# layout(mat);
par(mfrow=c(2,2))
for(i in 1:length(ex[1,]) ) { 
  hist(ex[,i], main=experimental_design$source_name_ch1[i]) 
}
# layout(1)
par(mfrow=c(1,1))
boxplot(ex)
```
All the expression data are centered around zero? No need for log2 transformation. 

According to GSE2627 summary. These data are log2 transformed expression ratios and are normalized by scaled loess normalization. 
```{r}
apply(ex, 2, mean)
apply(ex, 2, median)
```


Set up the data and proceed with analysis. Modified by GEO sample script. 
```{r, comment=NA}
gset@phenoData@data[,  c("source_name_ch1")]
gset@phenoData@data$source_name_ch1
treatments <- make.names(gset@phenoData@data$source_name_ch1)
fl.names = unique(treatments)
fl = as.factor( LETTERS[1:length(fl.names)]) 
names(fl) = fl.names
gset$description <- fl
design <- model.matrix(~ description + 0, gset)
colnames(design) <- levels(fl)
fit <- lmFit(gset, design)
cont.matrix <- makeContrasts( A-B, levels=fl)
fit2 <- contrasts.fit(fit, cont.matrix)
fit2 <- eBayes(fit2, 0.01)
tT <- topTable(fit2, adjust="fdr", sort.by="B", number=length(fit2))
```

```{r, comment=NA}
str(fit)
str(fit2)
str(tT)
```

Vocalno plot
```{r, comment=NA}

```



